---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
```

## Loading and preprocessing the data

```{r data}
df <- read.csv(unz("activity.zip", "activity.csv"))
df$date <- as.Date(df$date, format = "%Y-%m-%d")
df_complete <- df[complete.cases(df), ]
```

## What is the mean total number of steps taken per day?

Summary of the steps:
```{r summary}
summary(df_complete$steps)
```

Histogram of the total number of steps taken each day:
```{r histogram}
dfsum <- df_complete %>% group_by(date) %>% summarise_at("steps", sum, na.rm = TRUE)
hist(dfsum$steps, breaks = 30, xlab = "Number of steps per day", main = "")
```

Mean of the total number of steps taken per day:
```{r}
mean(dfsum$steps)
```

Median of the total number of steps taken per day:
```{r}
median(dfsum$steps)
```

## What is the average daily activity pattern?

```{r}
dfint <- df_complete %>% group_by(interval) %>% summarise_at("steps", mean, na.rm = TRUE)
ggplot(dfint, aes(interval, steps)) + labs(title = "Average daily activity pattern", y = "Average number of steps")  + geom_line()
```

5-minute interval with maximum average number of steps
```{r}
dfint[which.max(dfint$steps), ][[1]]
```

Average number of steps in that interval
```{r}
dfint[which.max(dfint$steps), ][[2]]
```

## Imputting missing values

Number of missing values: 
```{r}
sum(!complete.cases(df))
```

Filling a new dataframe with the average value for that 5-minute interval:
```{r}
df2 <- df

for (i in unique(df$interval)){
  df2[which(is.na(df$steps) & df$interval == i), 1] <- dfint[which(dfint$interval == i), 2]
}
```

Number of missing values with the new dataframe:
```{r}
sum(!complete.cases(df2))
```

Summary of the steps:
```{r}
summary(df2$steps)
```

Histogram of the total number of steps taken each day:
```{r}
df2sum <- df2 %>% group_by(date) %>% summarise_at("steps", sum, na.rm = TRUE)
hist(df2sum$steps, breaks = 30, xlab = "Number of steps per day", main = "")
```

Mean of the total number of steps taken per day:
```{r}
mean(df2sum$steps)
```

Median of the total number of steps taken per day:
```{r}
median(df2sum$steps)
```

## Are there differences in activity patterns between weekdays and weekends?

Creating type of day factor variable:
```{r}
df2["daytype"] = "weekday"
df2[which(weekdays(df2$date, abbreviate = TRUE) == "sá."), 4] <- "weekend"
df2[which(weekdays(df2$date, abbreviate = TRUE) == "do."), 4] <- "weekend"
df2$daytype <- as.factor(df2$daytype)
```

Creating comparison plot:
```{r}
df3 <- df2 %>% group_by(interval, daytype) %>% summarise_at("steps", mean, na.rm = TRUE)

ggplot(df3, aes(interval, steps)) + geom_line(aes(col = daytype)) + labs(title = "Activity pattern between weekdays and weekends", y = "Average number of steps") + facet_grid(rows = vars(daytype))
```
